---
title: "2: Sales Analysis"
date: today
execute:
  eval: false
---



## Executive Summary

*Write a 2-3 sentence summary of your analysis and key findings here after completing the assignment.*

---

## Introduction

As a financial analyst, I've been tasked with preparing Q1 2027 sales analysis for management reporting. The challenge is that our sales transaction data comes from one system, while customer master data comes from another system. Both datasets have common data quality issues that need to be resolved before analysis can begin.

In this blog post, I'll demonstrate a systematic approach to:

- Assess data quality
- Clean inconsistent formatting
- Join related datasets
- Create management-ready analysis
- Document the complete process for reproducibility

---

## Data Sources and Setup

### Required Libraries

```{r}
#| label: setup
#| message: false
#| warning: false
# Load required packages for data wrangling
library(tidyverse)  # For data manipulation and visualization
library(scales)     # For formatting numbers in output
```

### Creating Sample Data

Since this is a learning exercise, I'll create realistic sample datasets that mirror the types of data quality issues encountered in professional settings.

```{r}
#| label: create-data
# Sales Transactions Data (from Sales System)
# This data has various formatting issues typical of system exports
sales_raw <- tribble(
  ~transaction_id, ~customer_id, ~sale_date,   ~amount,      ~product_category, ~sales_rep,
  "TXN-001",       "CUST001",    "01/15/2027", "$28,500.00", "Hardware",        "Smith, John",
  "TXN-002",       "CUST002",    "01/20/2027", "7250",       "Software",        "johnson, mary",
  "TXN-003",       "CUST003",    "2027-02-01", "$18,000.50", "Software",        "Brown, Robert",
  "TXN-004",       "CUST001",    "02/15/2027", "22750.00",   "Services",        "Smith, John",
  "TXN-005",       "CUST999",    "02/28/2027", "$6,000",     "Hardware",        "Davis, Sarah",  # Invalid customer
  "TXN-006",       "CUST004",    "3/10/2027",  "14750",      "Hardware",        "wilson, david",
  "TXN-007",       "CUST002",    "03/15/2027", "$30,500",    "Services",        "johnson, mary",
  "TXN-008",       "CUST003",    "2027-03-20", "4250.00",    "Software",        "Brown, Robert"
)

# Customer Master Data (from CRM System)
# This data has inconsistent naming and missing values
customers_raw <- tribble(
  ~customer_id, ~customer_name,      ~region, ~segment,     ~credit_limit,
  "CUST001",    "acme corporation",  "North", "Enterprise", 100000,
  "CUST002",    "GLOBEX INC",        "South", "Mid-Market", 75000,
  "CUST003",    "umbrella corp",     "West",  "",           50000,         # Missing segment
  "CUST004",    "Stark Industries",  "East",  "Enterprise", 150000,
  "CUST005",    "wayne enterprises", "North", "Enterprise", 200000         # No sales in Q1
)

# Display raw data to see the issues
sales_raw
```

```{r}
#| label: show-customers
customers_raw
```

---

## Data Quality Assessment

Before cleaning, let's assess what issues we're dealing with:

```{r}
#| label: assess-sales-structure
# Examine sales data structure and issues
glimpse(sales_raw)
```

Let's look at the specific data quality issues in our sales data:

```{r}
#| label: assess-date-formats
# Date format analysis
sales_raw |>
  count(sale_date, sort = TRUE)
```

```{r}
#| label: assess-amount-formats
# Amount format analysis
sales_raw |>
  count(amount, sort = TRUE)
```

Now let's examine our customer master data:

```{r}
#| label: assess-customer-structure
# Examine customer data issues
glimpse(customers_raw)
```

```{r}
#| label: assess-customer-names
# Customer name format issues
customers_raw |>
  count(customer_name, sort = TRUE)
```

```{r}
#| label: check-orphaned-customers
# Check for potential orphaned sales (customers not in master data)
orphaned_customers <- sales_raw |>
  anti_join(customers_raw, by = "customer_id") |>
  distinct(customer_id)

orphaned_customers
```

Based on this assessment, we can see several data quality issues that need to be addressed:
- **Date formats**: Mix of MM/DD/YYYY and YYYY-MM-DD formats
- **Amount formatting**: Some amounts have dollar signs and commas, others don't
- **Name consistency**: Customer names and sales rep names have inconsistent capitalization
- **Missing data**: Some customer segments are blank
- **Orphaned records**: Customer ID "CUST999" appears in sales but not in customer master data

---

## Data Cleaning Process

### Step 1: Clean Sales Transaction Data

```{r}
#| label: clean-sales
sales_clean <- sales_raw |>
  mutate(
    # Fix date formatting - handle multiple date formats
    sale_date = case_when(
      str_detect(sale_date, "^\\d{1,2}/\\d{1,2}/\\d{4}$") ~ _____(sale_date),
      str_detect(sale_date, "^\\d{4}-\\d{2}-\\d{2}$") ~ _____(sale_date),
      TRUE ~ mdy(sale_date)  # safe fallback: NA instead of an error
    ),

    # Clean amount field - remove $ signs and commas, convert to numeric
    amount_clean = as.numeric(str_remove_all(amount, "[_____]")),

    # Standardize sales rep names to proper case
    sales_rep = str_to_title(_____),

    # Standardize product categories
    product_category = str_to_title(_____),

    # Add fiscal quarter for analysis
    fiscal_quarter = paste0("Q", quarter(_____)),

    # Add month name for reporting
    month_name = month(_____, label = TRUE, abbr = FALSE)
  ) |>
  # Remove original amount column, keep the clean version
  select(-amount) |>
  # Rename for clarity
  rename(amount = _____)

# Verify our cleaning worked
glimpse(sales_clean)
```

Let's verify our date conversion was successful:

```{r}
#| label: verify-dates
# Check date conversion success
sales_clean |>
  count(sale_date, month_name, fiscal_quarter)
```

### Step 2: Clean Customer Master Data

```{r}
#| label: clean-customers
customers_clean <- customers_raw |>
  mutate(
    # Standardize customer names to title case
    customer_name = _____(customer_name),

    # Fill missing segments with "Unclassified"
    segment = if_else(segment == "" | is.na(segment), "_____", segment),

    # Ensure region is properly formatted
    region = str_to_title(_____)
  )

# Verify customer data cleaning
customers_clean
```

Let's check that our segment cleaning worked properly:

```{r}
#| label: verify-segments
# Check segment cleaning
customers_clean |>
  count(segment)
```

The cleaning process has standardized our data formats and filled in missing values. All dates are now in proper Date format, amounts are numeric, and text fields have consistent capitalization.

---

## Data Integration

### Join Sales Transactions with Customer Data

```{r}
#| label: join-data
# Perform left join to keep all sales, add customer info where available
sales_with_customers <- sales_clean |>
  _____(customers_clean, by = "customer_id")

# Check the join results
glimpse(sales_with_customers)
```

Now let's identify any data quality issues from our join:

```{r}
#| label: check-missing-customers
# Identify sales without customer information (data quality issue)
missing_customer_info <- sales_with_customers |>
  filter(is.na(customer_name)) |>
  select(transaction_id, customer_id, amount, sale_date)

missing_customer_info
```

### Data Quality Report

```{r}
#| label: quality-report
# Create a data quality summary
data_quality_summary <- tibble(
  metric = c(
    "Total Sales Transactions",
    "Transactions with Valid Customers",
    "Transactions with Missing Customer Info",
    "Data Quality Rate"
  ),
  value = c(
    nrow(sales_with_customers),
    sum(!is.na(sales_with_customers$customer_name)),
    sum(is.na(sales_with_customers$customer_name)),
    paste0(round(sum(!is.na(sales_with_customers$customer_name)) /
                 nrow(sales_with_customers) * 100, 1), "%")
  )
)

data_quality_summary
```

Our data integration shows that we have one transaction (TXN-005) with an invalid customer ID that needs to be resolved before final reporting.

---

## Analysis and Insights

### Sales Performance by Month

```{r}
#| label: monthly-analysis
monthly_sales <- sales_with_customers |>
  # Only include transactions with valid customer data for analysis
  filter(!is.na(customer_name)) |>
  group_by(month_name) |>
  summarise(
    total_sales = sum(_____),
    transaction_count = n(),
    average_sale = mean(_____),
    .groups = "drop"
  ) |>
  # Calculate month-over-month change
  mutate(
    sales_change = _____ - lag(total_sales),
    sales_change_pct = (sales_change / lag(total_sales)) * 100
  )

monthly_sales
```

### Sales Performance by Region

```{r}
#| label: regional-analysis
regional_sales <- sales_with_customers |>
  filter(!is.na(customer_name)) |>
  group_by(_____) |>
  summarise(
    total_sales = _____,
    customer_count = n_distinct(customer_id),
    avg_sales_per_customer = total_sales / customer_count,
    .groups = "drop"
  ) |>
  arrange(desc(total_sales))

regional_sales
```

### Customer Segment Analysis

```{r}
#| label: segment-analysis
segment_performance <- sales_with_customers |>
  filter(!is.na(customer_name)) |>
  group_by(_____) |>
  summarise(
    total_sales = sum(amount),
    transaction_count = _____,
    customer_count = n_distinct(_____),
    avg_transaction_size = total_sales / transaction_count,
    .groups = "drop"
  ) |>
  # Calculate percentage of total sales
  mutate(
    sales_percentage = (total_sales / sum(total_sales)) * 100
  ) |>
  arrange(desc(total_sales))

segment_performance
```

### Top Performing Products

```{r}
#| label: product-analysis
product_performance <- sales_with_customers |>
  filter(!is.na(customer_name)) |>
  group_by(_____) |>
  summarise(
    total_sales = _____,
    transaction_count = n(),
    avg_sale_size = mean(amount),
    .groups = "drop"
  ) |>
  arrange(desc(_____))

product_performance
```

---

## Key Findings and Recommendations

### Summary of Findings

```{r}
#| label: summary-calcs
# Calculate key metrics for summary
total_q1_sales <- sales_with_customers |>
  filter(!is.na(customer_name)) |>
  summarise(total = sum(amount)) |>
  pull(total)

total_transactions <- sales_with_customers |>
  filter(!is.na(customer_name)) |>
  nrow()

data_quality_rate <- (sum(!is.na(sales_with_customers$customer_name)) /
                     nrow(sales_with_customers)) * 100

# Display formatted results
summary_metrics <- tibble(
  Metric = c("Total Q1 Sales", "Total Transactions", "Data Quality Rate"),
  Value = c(
    scales::dollar(total_q1_sales),
    scales::comma(total_transactions),
    paste0(round(data_quality_rate, 1), "%")
  )
)

summary_metrics
```

### Business Insights

*Based on the analysis above, write 3-4 key business insights. For example:*

1. **Sales Trend**: [Describe what you found in the monthly analysis - did sales increase or decrease month over month? What was the pattern?]

2. **Regional Performance**: [Summarize the regional findings - which region performed best? Are there opportunities in underperforming regions?]

3. **Customer Segments**: [Explain the segment analysis results - which segments drive the most revenue? What's the transaction pattern by segment?]

4. **Product Mix**: [Discuss product category performance - which products sell best? Are there cross-selling opportunities?]

### Data Quality Recommendations

1. **Customer Data Integration**: We identified one sales transaction with invalid customer ID "CUST999" that needs to be resolved with the CRM team. This represents _____ of our transactions that cannot be properly attributed.

2. **Data Standardization**: Implement consistent formatting standards across systems, particularly for:
   - Date formats (standardize on YYYY-MM-DD)
   - Currency fields (remove formatting, store as numeric)
   - Name fields (implement title case standards)

3. **Regular Data Quality Monitoring**: Establish automated checks for:
   - Orphaned transactions without valid customer references
   - Missing or blank required fields
   - Inconsistent formatting that could affect analysis

---

## Technical Methodology

### Data Wrangling Approach

The analysis followed a systematic data wrangling methodology:

1. **Assessment**: Used `glimpse()` and `count()` to understand data structure and quality issues
2. **Cleaning**: Applied `mutate()` with string functions and date conversion to standardize formats
3. **Integration**: Used `left_join()` to combine datasets while preserving all sales data
4. **Quality Control**: Used `anti_join()` to identify orphaned records and `filter()` with `is.na()` to check join results
5. **Analysis**: Applied `group_by()` and `summarise()` to create business insights

### Key Functions and Techniques Used

- **Date Handling**: `mdy()`, `ymd()`, `quarter()`, `month()` for standardizing date formats
- **String Cleaning**: `str_remove_all()`, `str_to_title()` for text standardization
- **Data Integration**: `left_join()`, `anti_join()` for combining datasets and quality checking
- **Conditional Logic**: `case_when()`, `if_else()` for handling multiple scenarios
- **Data Aggregation**: `group_by()`, `summarise()`, `n()`, `n_distinct()` for business analysis

