---
title: "4: Data Visualization for Accounting"
date: today
execute:
  eval: false
  message: false
  warning: false
---



## Executive Summary

*Write a 2-3 sentence summary of your visualizations and key insights here after completing the assignment.*

---

## Introduction

As an accounting professional, you need to communicate complex financial information clearly. Effective data visualization helps stakeholders understand:

- Financial performance trends
- Budget variances and anomalies
- Departmental spending patterns
- Revenue growth across customer segments

In this blog post, you will:

- Create professional bar charts for categorical comparisons
- Build line charts to show financial trends
- Design dashboard layouts for comprehensive reporting
- Apply best practices for ethical data visualization

## Setup and Load Libraries

### Required Libraries

```{r}
#| label: setup
# Load required packages
library(tidyverse)  # For ggplot2 and data manipulation
library(scales)     # For formatting currency and percentages
library(patchwork)  # For combining multiple plots
```

### Creating Sample Financial Data

Let's create sample data representing a company's financial performance:

**Monthly Financial Performance**

```{r}
#| label: monthly-data
# Monthly financial data for 2027
monthly_financials <- tribble(
  ~month,       ~revenue, ~expenses, ~net_income, ~quarter,
  "2027-01-01", 131000,   90000,     41000,       "Q1",
  "2027-02-01", 135000,   100000,    35000,       "Q1",
  "2027-03-01", 141000,   99000,     42000,       "Q1",
  "2027-04-01", 148000,   105000,    43000,       "Q2",
  "2027-05-01", 151000,   106000,    45000,       "Q2",
  "2027-06-01", 165000,   120000,    45000,       "Q2",
  "2027-07-01", 173000,   118000,    55000,       "Q3",
  "2027-08-01", 186000,   141000,    45000,       "Q3",
  "2027-09-01", 200000,   150000,    50000,       "Q3",
  "2027-10-01", 211000,   152000,    59000,       "Q4",
  "2027-11-01", 215000,   152000,    63000,       "Q4",
  "2027-12-01", 225000,   168000,    57000,       "Q4"
)
monthly_financials <- monthly_financials |>
  mutate(month = as.Date(month))

# View the data
monthly_financials
```

**Department Expense Data**

```{r}
#| label: dept-data
# Department expense breakdown
dept_expenses <- tribble(
  ~department,  ~expense_category, ~amount,
  "Sales",      "Salaries",        224548,
  "Sales",      "Travel",          78573,
  "Sales",      "Supplies",        32128,
  "Marketing",  "Salaries",        149358,
  "Marketing",  "Advertising",     122923,
  "Marketing",  "Supplies",        25735,
  "Operations", "Salaries",        324254,
  "Operations", "Equipment",       106296,
  "Operations", "Supplies",        68886,
  "Finance",    "Salaries",        188399,
  "Finance",    "Software",        47779,
  "Finance",    "Supplies",        27768
)

# View the data
dept_expenses
```

### Setting Professional Defaults

```{r}
#| label: defaults
# Set a clean theme for all plots
theme_set(theme_minimal())

# Prevent scientific notation in output
options(scipen = 9999)
```

## Creating Your First Visualization

### Bar Chart: Department Total Expenses

```{r}
#| label: dept-bar
# Calculate total expenses by department
dept_totals <- dept_expenses |>
  group_by(_____) |>  # Group by department
  summarize(total_expenses = _____) |>  # Sum the amounts
  arrange(desc(total_expenses))

# Display dept_totals
dept_totals

# Create a horizontal bar chart
ggplot(dept_totals, aes(x = _____, y = reorder(department, _____))) +
  geom_col(fill = "dark blue") +
  scale_x_continuous(labels = _____) +  # Format as currency
  labs(
    title = "Total Expenses by Department",
    subtitle = "Fiscal Year 2027",
    x = NULL,
    y = NULL,
    caption = "Source: General Ledger"
  )
```

### Line Chart: Revenue Trend Analysis

```{r}
#| label: revenue-line
# Create a line chart showing monthly revenue
monthly_financials |>
  ggplot(aes(x = _____, y = _____)) +  # Map month to x, revenue to y
  geom_line(color = "dark green", size = 1.2) +
  geom_point(color = "dark green", size = 2.5) +
  scale_x_date(
    labels = label_date_short(),
    breaks = breaks_width("1 months")) +
  scale_y_continuous(labels = _____) +  # Format as currency
  labs(
    title = "_____",  # Add appropriate title
    subtitle = "January - December 2027",
    x = NULL,
    y = NULL,
    caption = "Source: Financial Reporting System"
  )
```

## Multi-Series Line Chart

### Comparing Revenue vs Expenses

```{r}
#| label: multi-line
# Transform data for multi-line plotting
financial_trends <- monthly_financials |>
  select(month, revenue, expenses) |>
  pivot_longer(
    cols = c(_____, _____),  # Select revenue and expenses columns
    names_to = "metric",
    values_to = "amount"
  )


# Display financial_trends
_____


# Create multi-line chart
ggplot(financial_trends, aes(x = month, y = amount, color = _____)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  scale_x_date(
    labels = label_date_short(),
    breaks = breaks_width("3 months")) +
  scale_y_continuous(labels = dollar_format()) +
  scale_color_manual(
    values = c("revenue" = "_____", "expenses" = "_____"),  # Choose appropriate colors
    labels = c("revenue" = "Revenue", "expenses" = "Expenses"),
    breaks = c("revenue", "expenses") # Set order of legend items
  ) +
  labs(
    title = "Revenue vs Expenses Trend",
    subtitle = "2027 Financial Performance",
    x = NULL,
    y = NULL,
    color = NULL
  ) +
  theme(legend.position = "bottom")
```

## Department Expense Breakdown

### Stacked Bar Chart by Category

```{r}
#| label: stacked-bar
# Create stacked bar chart showing expense categories
dept_expenses |>
  ggplot(aes(x = department, y = amount, fill = _____)) +  # Fill by expense_category
  geom_col(position = "_____") +  # Use stack position
  scale_y_continuous(labels = dollar_format()) +
  scale_fill_brewer(palette = "Blues") +
  labs(
    title = "Department Expenses by Category",
    subtitle = "Fiscal Year 2027",
    x = NULL,
    y = NULL,
    fill = NULL,
    caption = "Source: General Ledger"
  ) +
  theme(legend.position = "bottom")
```

## Creating a Financial Dashboard

### Panel 1: Net Income Trend

```{r}
#| label: panel-net-income
# Create net income visualization
p1 <- monthly_financials |>
  ggplot(aes(x = month, y = _____)) +  # Plot net_income
  geom_col(fill = ifelse(monthly_financials$net_income > 0, "seagreen", "tomato")) +
  scale_x_date(labels = label_date_short()) +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    title = "Monthly Net Income",
    x = NULL,
    y = NULL
  )

# Display p1
_____
```

### Panel 2: Quarterly Performance

```{r}
#| label: panel-quarterly
# Summarize by quarter
quarterly_summary <- monthly_financials |>
  group_by(_____) |>  # Group by quarter
  summarize(
    total_revenue = sum(_____),
    total_expenses = sum(_____),
    total_net_income = sum(_____)
  )

# Display quarterly_summary
_____

# Create quarterly comparison
p2 <- quarterly_summary |>
  ggplot(aes(x = _____)) +
  geom_col(aes(y = total_revenue), fill = "seagreen") +
  geom_col(aes(y = -total_expenses), fill = "tomato") +
  scale_y_continuous(
    labels = function(x) dollar_format()(abs(x))
  ) +
  labs(
    title = "Quarterly Revenue vs Expenses",
    subtitle = "Revenue (green) above, Expenses (red) below",
    x = NULL,
    y = NULL
  )


# Display p2
p2

```

### Combine Dashboard Panels

```{r}
#| label: dashboard
# Combine panels using patchwork
dashboard <- p1 / p2

# Display dashboard
dashboard +
  plot_annotation(
    title = "Financial Performance Dashboard",
    subtitle = "Fiscal Year 2027",
    caption = "Source: Financial Reporting System"
  )
```

## Advanced Visualization: Variance Analysis

```{r}
#| label: variance
# Create budget vs actual data
budget_data <- tribble(
  ~department,  ~budget, ~actual,
  "Sales",      350000,  335249,
  "Marketing",  320000,  298016,
  "Operations", 480000,  499436,
  "Finance",    240000,  263946
)
budget_data <- budget_data |>
  mutate(
    variance_amount = _____ - _____,  # Calculate variance
    variance_pct = (variance_amount / budget) * 100,
    favorable = variance_amount < 0  # For expenses, under budget is favorable
  )


# Display budget_data
_____


# Create variance visualization
budget_data |>
  ggplot(aes(x = variance_pct, y = reorder(department, variance_pct))) +
  geom_col(aes(fill = favorable)) +
  scale_fill_manual(
    values = c("TRUE" = "_____", "FALSE" = "_____"),  # Green for favorable, red for unfavorable
    labels = c("TRUE" = "Favorable", "FALSE" = "Unfavorable")
  ) +
  scale_x_continuous(labels = percent_format(scale = 1)) +
  labs(
    title = "Budget Variance by Department",
    subtitle = "Percentage Over/Under Budget",
    x = NULL,
    y = NULL,
    fill = NULL
  ) +
  theme(legend.position = "bottom")
```

## Key Insights and Best Practices

### Data Visualization Ethics

```{r}
#| label: ethics
# Example of ethical vs misleading visualization
# Create the same data with different y-axis scales

# Misleading (truncated y-axis)
misleading <- monthly_financials |>
  ggplot(aes(x = month, y = revenue)) +
  geom_line(color = "dark green", size = 1.2) +
  scale_y_continuous(
    limits = c(min(monthly_financials$revenue) - 5000,
               max(monthly_financials$revenue) + 10000),  # Truncated axis
    labels = dollar_format()
  ) +
  labs(title = "MASSIVE Revenue Growth!")

# Ethical (starts at zero)
ethical <- monthly_financials |>
  ggplot(aes(x = month, y = revenue)) +
  geom_line(color = "dark green", size = 1.2) +
  scale_y_continuous(
    limits = c(0, NA),  # Start at zero
    labels = dollar_format()
  ) +
  labs(title = "Monthly Revenue Growth")

# Compare side by side
misleading | ethical
```

### Summary Metrics

```{r}
#| label: summary-metrics
# Calculate key performance metrics
performance_summary <- monthly_financials |>
  summarize(
    total_revenue = sum(revenue),
    total_expenses = sum(expenses),
    total_net_income = sum(net_income),
    profit_margin = (total_net_income / total_revenue),
    revenue_growth = ((last(revenue) - first(revenue)) / first(revenue))
  )


# Display performance_summary
performance_summary


# Format the summary
performance_summary |>
  mutate(
    total_revenue = dollar(total_revenue),
    total_expenses = dollar(total_expenses),
    total_net_income = dollar(total_net_income),
    profit_margin = percent(profit_margin),
    revenue_growth = percent(revenue_growth)
  )
```

## Key Findings

Based on the visualizations above, complete these insights:

- **Revenue Trend**: Revenue (increased/decreased) by _____ from January to December.
- **Highest Spending Department**: _____ had the highest total expenses at _____.
- **Profit Margin**: The company achieved a _____ profit margin for the year.
- **Best Performing Quarter**: _____ showed the highest net income at _____.
- **Budget Performance**: The _____ and _____ departments were under budget (favorable variance).

