---
title: "6: Clustering Analysis for Audit Risk Detection"
date: today
execute: 
  eval: false #set to true
  message: false
  warning: false
---

## Executive Summary

*Write a 2-3 sentence summary of your clustering analysis and the highest-risk expense patterns identified. Complete this section after finishing the assignment.*

---

## Introduction

As an accounting professional performing audit procedures, you need to identify unusual patterns and potential anomalies in expense data. Clustering analysis helps auditors:

- Group similar transactions to identify outliers
- Detect expense patterns that deviate from normal behavior
- Prioritize high-risk transactions for detailed review
- Improve audit efficiency through data-driven sampling

In this blog post, you will:

- Apply k-means clustering to expense report data
- Visualize clusters to identify unusual patterns
- Analyze cluster characteristics for audit risk indicators
- Recommend transactions for further investigation

## Setup and Load Libraries

### Required Libraries

```{r}
#| label: setup
# Load packages we need - matching the slides
library(tidyverse)    # For data manipulation
library(tidymodels)   # For clustering
library(tidyclust)    # Additional clustering tools
library(scales)       # For formatting numbers
library(gt)           # For nice tables
library(patchwork)    # For combining multiple plots

# Set preferences
theme_set(theme_minimal())  # Clean plots
options(scipen = 999)       # No scientific notation
set.seed(2027)              # Reproducible results
```

### Load and Explore the Data

```{r}
# Load the expense report data
expense_reports <- read_csv("https://estanny.org/data/06_expense_reports.csv")

# Display structure of the data
glimpse(expense_reports)

# Basic summary statistics
summary(expense_reports)
```

## Initial Data Exploration

### Expense Distribution Analysis

```{r}
# Visualize expense amount distribution
expense_reports |>
  ggplot(aes(x = amount)) +
  geom_histogram(bins = 50, fill = "steelblue") +
  scale_x_continuous(labels = dollar_format()) +
  labs(
    title = "Distribution of Expense Amounts",
    subtitle = "Most expenses cluster at lower amounts with some outliers",
    x = "Expense Amount",
    y = "Count"
  )

# Check for potential red flags
red_flags_summary <- expense_reports |>
  summarize(
    total_expenses = n(),
    weekend_expenses = sum(_____),  # Count weekend expenses
    round_amounts = sum(_____),     # Count round amounts
    high_amounts = sum(_____),      # Count high amounts
    delayed_submissions = sum(_____) # Count delayed submissions
  )

# Display red flags summary using gt
red_flags_summary |>
  pivot_longer(everything(), names_to = "Metric", values_to = "Count") |>
  gt() |>
  tab_header(title = "Expense Report Red Flags Summary")
```

### Expense Patterns by Department

```{r}
# Summarize by department
dept_summary <- expense_reports |>
  group_by(_____) |>  # Group by department
  summarize(
    expense_count = n(),
    total_amount = sum(_____),
    avg_amount = mean(_____),
    weekend_pct = mean(_____),
    round_amount_pct = mean(_____) 
  ) |>
  arrange(desc(total_amount))

# Display department summary using gt
dept_summary |>
  mutate(
    total_amount = dollar(total_amount, accuracy = 1),
    avg_amount = dollar(avg_amount, accuracy =1),
    weekend_pct = percent(weekend_pct, accuracy = 1),
    round_amount_pct = percent(round_amount_pct, accuracy = 1)
  ) |>
  gt() |>
  tab_header(title = "Department Expense Analysis")

# Visualize department spending
dept_summary |>
  ggplot(aes(x = total_amount, y = reorder(department, total_amount))) +
  geom_col(fill = "darkblue") +
  scale_x_continuous(labels = _____) +  # Format as currency
  labs(
    title = "Total Expenses by Department",
    x = NULL,
    y = NULL
  )
```

## Data Preparation 

Prepare the data for clustering by selecting and scaling variables (following the slides approach).

```{r}
# Select features for clustering
# Document why you chose these features
clustering_features <- expense_reports |>
  select(
    _____,                    # Transaction amount - consider log transform
    _____,                    # Days to submit
    _____,                    # Weekend indicator (0/1)
    _____,                    # Round amount flag (0/1)
    _____,                    # Multiple expenses same day
    _____                     # Employee's average expense
  )

# Add log transformation for amount if needed
clustering_features <- clustering_features |>
  mutate(log_amount = log(amount + 1)) |>  # Add 1 to handle any zeros
  select(-amount)  # Remove original amount

# Scale the features (following slides approach)
clustering_data <- clustering_features |>
  mutate(
    log_amount_scaled = scale(log_amount)[,1],
    submission_delay_scaled = scale(_____)[,1],
    is_weekend_scaled = _____,               # Binary - keep as is
    is_round_amount_scaled = _____,          # Binary - keep as is
    same_day_expense_count_scaled = scale(_____)[,1],
    employee_avg_amount_scaled = scale(_____)[,1]
  ) |>
  select(ends_with("_scaled"))

# Check your work
clustering_data |> summary()
```

## K-means Clustering Analysis

### Step 1: Set Up K-Means Model

```{r}
# Specify the k-means model with 3 clusters (following slides)
kmeans_spec <- k_means(num_clusters = _____) |>  # Use 3 clusters
  set_engine("stats")

# View the model specification
kmeans_spec
```

### Step 2: Fit the Model

```{r}
# Fit the model to our prepared data
kmeans_fit <- kmeans_spec |>
  fit(~ ., data = _____)  # Fit to all variables in clustering_data

```

### Step 3: Extract Cluster Assignments

```{r}
# Get cluster assignments
cluster_assignments <- kmeans_fit |>
  extract_cluster_assignment()

# View the assignments
head(cluster_assignments)

# Add clusters to original data
expense_reports_clustered <- expense_reports |>
  bind_cols(cluster_assignments)

# Check the results
expense_reports_clustered |>
  select(expense_id, amount, submission_delay, is_weekend, .cluster) |>
  head(10)
```

## Visualizing Clusters

### Visualization 1: Amount vs Submission Delay

```{r}
# Create scatter plot colored by cluster
expense_reports_clustered |>
  ggplot(aes(x = submission_delay, y = amount, color = .cluster)) +
  geom_point(alpha = 0.6, size = 2) +
  scale_y_continuous(labels = dollar_format()) +
  scale_color_manual(
    values = c("Cluster_1" = "orange", "Cluster_2" = "red", "Cluster_3" = "green")
  ) +
  labs(
    title = "Expense Clusters: Amount vs Submission Delay",
    x = "Days to Submit",
    y = "Expense Amount",
    color = NULL 
  )
```

### Visualization 2: Box Plots by Cluster

```{r}
# Create box plots of amount by cluster
expense_reports_clustered |>
  ggplot(aes(x = .cluster, y = amount, fill = .cluster)) +
  geom______() +
  scale_y_continuous(labels = dollar_format()) +
  scale_fill_manual(
    values = c("Cluster_1" = "orange", "Cluster_2" = "red", "Cluster_3" = "green")
  ) +
  labs(
    title = "Expense Amount Distribution by Cluster",
    x = "Cluster",
    y = "Amount"
  ) +
  theme(legend.position = "none")
```

## Cluster Analysis and Risk Assessment

### Step 4: Analyze Cluster Characteristics

```{r}
# Calculate detailed statistics for each cluster
cluster_profiles <- expense_reports_clustered |>
  group_by(_____) |>  # Group by cluster
  summarize(
    transaction_count = n(),
    avg_amount = mean(amount),
    max_amount = max(amount),
    avg_submission_delay = mean(submission_delay),
    weekend_pct = mean(is_weekend),
    round_amount_pct = mean(is_round_amount),
    high_amount_pct = mean(is_high_amount),
    multiple_same_day_pct = mean(multiple_same_day),
    unique_employees = n_distinct(employee_id),
    unique_vendors = n_distinct(vendor)
  )

# Display cluster profiles using gt
cluster_profiles |>
  mutate(
    avg_amount = dollar(avg_amount),
    max_amount = dollar(max_amount),
    weekend_pct = percent(weekend_pct),
    round_amount_pct = percent(round_amount_pct),
    high_amount_pct = percent(high_amount_pct),
    multiple_same_day_pct = percent(multiple_same_day_pct)
  ) |>
  gt() |>
  tab_header(title = "Cluster Characteristics Analysis")
```

### Step 5: Identify High-Risk Cluster

```{r}
# Add risk scoring (following slides approach)
cluster_profiles <- cluster_profiles |>
  mutate(
    risk_score = (avg_amount / max(avg_amount)) * _____ +     # Amount (40% weight)
                 ((50 - avg_submission_delay) / 50) * _____ + # Speed (30% weight)  
                 weekend_pct * _____ +                        # Weekend (20% weight)
                 round_amount_pct * _____                     # Round amounts (10% weight)
  ) |>
  arrange(desc(risk_score)) |>
  mutate(risk_level = c("High Risk", "Medium Risk", "Low Risk"))

# Display risk-scored clusters
cluster_profiles |>
  select(.cluster, risk_score, risk_level, transaction_count, avg_amount) |>
  mutate(avg_amount = dollar(avg_amount)) |>
  gt() |>
  tab_header(title = "Cluster Risk Assessment")

# Get highest risk cluster
high_risk_cluster <- cluster_profiles |>
  filter(risk_level == "High Risk") |>
  pull(.cluster)
```

### Step 6: Apply Risk Classifications

```{r}
# Create risk mapping
cluster_risk_mapping <- cluster_profiles |>
  select(.cluster, risk_level)

# Apply risk labels to all transactions
expense_reports_clustered <- expense_reports_clustered |>
  left_join(cluster_risk_mapping, by = "i.cluster")

# View high-risk transactions
high_risk_expenses <- expense_reports_clustered |>
  filter(risk_level == "High Risk") |>
  arrange(desc(amount))

# Display top 10 high-risk expenses
high_risk_expenses |>
  select(expense_id, employee_id, department, vendor, amount, 
        expense_date, is_weekend, is_round_amount) |>
  head(10) |>
  mutate(amount = dollar(amount)) |>
  gt() |>
  tab_header(title = "Top 10 High-Risk Expenses")
```

## Audit Planning 📝

Use clustering results to create an audit plan.

```{r}
#| eval: false

# Analyze high-risk cluster for audit planning
audit_focus <- expense_reports_clustered |>
  filter(.cluster == _____) |>  # Use high_risk_cluster
  summarise(
    transaction_count = _____,
    total_value = sum(_____),
    vendors_affected = n_distinct(_____),
    departments_affected = n_distinct(_____),
    avg_amount = mean(_____)
  )

# Format and display results
audit_focus |>
  mutate(
    total_value = dollar(total_value),
    avg_amount = dollar(avg_amount)
  ) |>
  gt() |>
  tab_header(title = "High-Risk Cluster Analysis for Audit Planning")

# Create audit recommendations
audit_recommendations <- tribble(
  ~Area, ~Finding, ~Recommendation,
  "Scope", paste(_____, "high-risk transactions identified"), 
  "Perform detailed testing on all high-risk transactions",
  "Materiality", paste("Total value:", dollar(_____)),
  "Focus on transactions over $_____",
  "Vendors", paste(_____, "vendors in high-risk cluster"),
  "Review vendor approval and payment processes"
)

audit_recommendations |>
  gt() |>
  tab_header(title = "Audit Planning Recommendations")
```

## Visualization Summary Dashboard

```{r}
# Create summary metrics following slides approach
library(patchwork)

# Risk indicators by cluster  
p1 <- cluster_profiles |>
  select(.cluster, weekend_pct, round_amount_pct, high_amount_pct) |>
  pivot_longer(-.cluster, names_to = "indicator", values_to = "percentage") |>
  mutate(
    indicator = case_when(
      indicator == "weekend_pct" ~ "Weekend",
      indicator == "round_amount_pct" ~ "Round Amount", 
      indicator == "high_amount_pct" ~ "High Amount"
    )
  ) |>
  ggplot(aes(x = indicator, y = percentage, fill = .cluster)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c("Cluster_1" = "green", "Cluster_2" = "orange", "Cluster_3" = "red")
  ) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Risk Indicators by Cluster",
    x = NULL,
    y = "Percentage",
    fill = "Cluster"
  )

# Department distribution in high-risk cluster
p2 <- expense_reports_clustered |>
  filter(.cluster == high_risk_cluster) |>
  count(department) |>
  ggplot(aes(x = reorder(department, n), y = n)) +
  geom_col(fill = "darkred") +
  coord_flip() +
  labs(
    title = "High-Risk Expenses by Department",
    x = NULL,
    y = "Count"
  )

# Combine plots
p1 / p2
```

## Key Findings and Recommendations

Based on the clustering analysis, complete these insights:

1. **Cluster Identification**: Cluster _____ was identified as the highest risk with _____ expenses totaling $_____.

2. **Risk Indicators**: This cluster showed:

- _____% weekend expenses (vs. _____% overall average)
- _____% round dollar amounts (potential red flag)
- Average submission delay of _____ days
- _____% of expenses from same-day multiple submissions

3. **Department Focus**: The _____ and _____ departments had the most high-risk expenses.

4. **Audit Recommendations**:

- Perform detailed review of all expenses over $_____ in the high-risk cluster
- Investigate employees with multiple same-day expenses
- Review weekend expense policies, especially for _____ department
- Implement controls for expenses with submission delays over _____ days

5. **Estimated Audit Coverage**: By focusing on the high-risk cluster, auditors can review _____% of expenses by count while covering _____% of total expense value.

