---
title: "10: Accounts Receivable Dashboard" 
format: dashboard
execute: 
  eval: false
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false

# Load required packages
library(tidyverse)   # for data manipulation
library(scales)      # for formatting numbers
library(DT)          # for interactive tables
library(plotly)      # for interactive plots
library(gt)          # for creating beautiful tables

# Set theme for consistent plot appearance
theme_set(theme_minimal(base_size = 12))

# Define color scheme for AR status
ar_colors <- c(
  current = "#2E7D32",      # Green for current
  early = "#FFC107",        # Yellow for 1-30 days
  moderate = "#FF9800",     # Orange for 31-60 days
  late = "#F44336",         # Red for 61-90 days
  critical = "#B71C1C",     # Dark red for over 90 days
  positive = "#1976D2",     # Blue for positive indicators
  negative = "#C62828"      # Red for negative indicators
)

# Load the AR data
ar_data  <- read_rds("https://estanny.org/data/10-assignment-ar_data.rds")

```

# Executive Summary {.tabset}

As of January 31, 2026, the company faces significant collection risk with $_______M (_______%) of the $_______M total accounts receivable past due, including $_______ severely delinquent beyond 90 days concentrated primarily in the _______ and _______ sectors. _______ of ten major customers are classified as high-risk, with credit utilization rates ranging from _______% to _______% above their limits—most notably the customers, _______ and _______—signaling an urgent need to reassess credit policies and enforce collection protocols on these accounts. Collection trends reveal an average days-to-pay of _______ days against standard 30-day terms, with DSO of _______ days indicating deteriorating collection efficiency that requires immediate intervention to prevent further aging. Based on historical collection patterns and current aging the cash collection forecast projects approximately $_______M in expected receipts and $_______K is at write-off risk.

## Key Metrics

### Row {height="25%"}

```{r}
#| label: load-data
# Load the AR data (using created sample data)
# ar_data <- read_rds("data/10-assignment-ar_data.rds")

# Set current date for aging calculations
current_date <- ymd("2026-01-31")

```

```{r}
#| label: calculate-kpis
# Calculate key performance indicators

# Total AR balance
total_ar <- ar_data |>
  _______(status == "Open") |>
  summarise(total = sum(_______)) |>
  pull(total)

# Calculate Days Sales Outstanding (DSO)
# Calculate average daily sales from last 90 days of paid invoices
avg_daily_sales <- ar_data |>
  filter(status == "Paid" & payment_date >= today() - days(90)) |>
  _______(daily_sales = sum(amount) / 90) |>
  pull(daily_sales)

current_dso <- round(total_ar / avg_daily_sales, 1)

# Calculate past due percentage
past_due_amount <- ar_data |>
  filter(status == "Open" & due_date < current_date) |>
  summarise(total = sum(amount)) |>
  pull(total)

past_due_percentage <- _______ / total_ar

# Count of overdue invoices
overdue_count <- ar_data |>
  filter(status == "Open" & due_date < current_date) |>
  nrow()

# Average days to pay (for paid invoices)
avg_days_to_pay <- ar_data |>
  _______(status == "Paid" & !is.na(payment_date)) |>
  _______(days_to_pay = as.numeric(payment_date - invoice_date)) |>
  _______(avg_days = mean(days_to_pay, na.rm = TRUE)) |>
  pull(avg_days)
```

### Row {height="75%"}

## Summary Statistics

```{r}
#| label: summary-table
# Create comprehensive AR summary statistics with Key Metrics
summary_stats <- tibble(
  Metric = c(
    "Days Sales Outstanding (DSO)",
    "Past Due %",
    "Overdue Invoice Count",
    "Avg Days to Pay",
    "Total Outstanding AR",
    "Number of Open Invoices",
    "Average Invoice Amount",
    "Median Invoice Amount",
    "Largest Invoice",
    "Smallest Invoice",
    "Total Past Due",
    "Current (Not Due)"
  ),
  Value = c(
    current_dso,
    past_due_percentage,
    overdue_count,
    avg_days_to_pay,
    total_ar,
    ar_data |> filter(status == "Open") |> nrow(),
    ar_data |> filter(status == "Open") |> summarise(mean(amount)) |> pull(),
    ar_data |> filter(status == "Open") |> summarise(median(amount)) |> pull(),
    ar_data |> filter(status == "Open") |> summarise(max(amount)) |> pull(),
    ar_data |> filter(status == "Open") |> summarise(min(amount)) |> pull(),
    past_due_amount,
    total_ar - past_due_amount
  ),
  Format = c("days", "percent", "number", "days",
             "currency", "number", "currency", "currency",
             "currency", "currency", "currency", "currency")
)

# Create summary table using gt
summary_stats |>
  select(Metric, Value) |>
  gt() |>
  _______(
    title = "Accounts Receivable Summary",
    subtitle = paste("As of", format(current_date, "%B %d, %Y"))
  ) |>
  tab_row_group(
    label = "Key Performance Indicators",
    rows = 1:4
  ) |>
  tab_row_group(
    label = "AR Details",
    rows = 5:12
  ) |>
  fmt_currency(
    columns = Value,
    rows = c(5, 7:12),
    decimals = 0
  ) |>
  fmt_number(
    columns = Value,
    rows = c(1, 4, 6),
    decimals = 1
  ) |>
  fmt_percent(
    columns = Value,
    rows = 2,
    decimals = 1
  ) |>
  fmt_number(
    columns = Value,
    rows = 3,
    decimals = 0
  ) |>
  tab_style(
    style = cell_fill(color = "#E3F2FD"),
    locations = cells_body(rows = c(1:4))
  ) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(rows = 11)
  )
```

# Aging Analysis

## Row {height="30%"}

### Aging Summary

```{r}
#| label: create-aging-buckets
# Create aging buckets for AR analysis
ar_aging <- ar_data |>
  _______status == "Open") |>
  _______(
    # Calculate days outstanding
    days_outstanding = as.numeric(current_date - invoice_date),
    # Calculate days overdue
    days_overdue = as.numeric(current_date - due_date),
    days_overdue = if_else(days_overdue < 0, 0, days_overdue),
    # Create aging buckets
    aging_bucket = case_when(
      days_overdue == 0 ~ "Current",
      days_overdue <= 30 ~ "1-30 Days",
      days_overdue <= 60 ~ "31-60 Days",
      days_overdue <= 90 ~ "61-90 Days",
      TRUE ~ "Over 90 Days"
    ),
    # Set factor levels for proper ordering
    aging_bucket = factor(aging_bucket, 
                         levels = c("Current", "1-30 Days", "31-60 Days", 
                                   "61-90 Days", "Over 90 Days"))
  )

# Create aging summary table
aging_summary <- ar_aging |>
  group_by(aging_bucket) |>
  summarise(
    `Invoice Count` = n(),
    `Total Amount` = sum(amount),
    `Percentage` = sum(amount) / sum(ar_aging$amount),
    .groups = "drop"
  )

# Display aging summary with gt
aging_summary |>
  gt() |>
  tab_header(
    title = "AR Aging Summary",
    subtitle = "Distribution by aging bucket"
  ) |>
  fmt_number(
    columns = `Invoice Count`,
    decimals = 0
  ) |>
  fmt_currency(
    columns = `Total Amount`,
    decimals = 0
  ) |>
  fmt_percent(
    columns = Percentage,
    decimals = 0
  ) |>
  cols_label(
    aging_bucket = "Aging Bucket"
  ) |>
  # Apply conditional formatting based on aging
  tab_style(
    style = cell_fill(color = "#C8E6C9"),
    locations = cells_body(
      columns = everything(),
      rows = aging_bucket == "Current"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(
      columns = everything(),
      rows = aging_bucket == "Over 90 Days"
    )
  )
```

## Row {height="35%"}


#### Aging Distribution

```{r}
#| label: aging-distribution-plot
# Create aging distribution bar chart
aging_plot <- ar_aging |>
  _______(aging_bucket) |>
  summarise(total = sum(amount), .groups = "drop") |>
  ggplot(aes(x = aging_bucket, y = total, fill = aging_bucket)) +
  geom_col() +
  scale_y_continuous(labels = dollar_format()) +
  scale_fill_manual(
    values = c(
      "Current" = ar_colors[["current"]],
      "1-30 Days" = ar_colors[["early"]],
      "31-60 Days" = ar_colors[["moderate"]],
      "61-90 Days" = ar_colors[["late"]],
      "Over 90 Days" = ar_colors[["critical"]]
    )
  ) +
  labs(
    title = "AR Aging Distribution",
    x = NULL,
    y = "Amount Outstanding",
    fill = NULL
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# Convert to interactive plot
_______(aging_plot, tooltip = c("x", "y")) |>
  layout(hovermode = "x unified")
```

## Row {height="35%"}

#### Aging by Customer Type

```{r}
#| label: aging-by-customer-type
# Create aging analysis by customer industry/type
industry_aging <- ar_aging |>
  _______(industry, aging_bucket) |>
  summarise(total = sum(amount), .groups = "drop")

# Create stacked bar chart 
industry_plot <- industry_aging |>
  ggplot(aes(y = reorder(industry, total, sum), x = total, fill = aging_bucket)) +  # Swapped x and y, added sum function
  geom_col(position = "stack") +
  scale_x_continuous(labels = dollar_format()) +  # Changed to scale_x
  scale_fill_manual(
    values = c(
      "Current" = ar_colors[["current"]],
      "1-30 Days" = ar_colors[["early"]],
      "31-60 Days" = ar_colors[["moderate"]],
      "61-90 Days" = ar_colors[["late"]],
      "Over 90 Days" = ar_colors[["critical"]]
    )
  ) +
  labs(
    title = "Aging by Industry",
    y = NULL,  # Swapped labels
    x = "Amount Outstanding",
    fill = NULL
  )

# Convert to interactive
ggplotly(_______) |>
  layout(hovermode = "y unified")  # Changed to y unified
```

# Customer Analysis

## Row

### Top Customers by Exposure

```{r}
#| label: top-customers-table
# Calculate customer exposure metrics
customer_summary <- ar_aging |>
  group_by(customer_id, customer_name, credit_limit) |>
  _______(
    total_ar = sum(amount),
    invoice_count = n(),
    oldest_invoice = min(invoice_date),
    days_oldest = as.numeric(current_date - oldest_invoice),
    avg_days_overdue = round(mean(days_overdue, na.rm = TRUE), 0),
    .groups = "drop"
  ) |>
  mutate(
    credit_utilization = total_ar / credit_limit,
    risk_score = case_when(
      avg_days_overdue > 60 | credit_utilization > 0.9 ~ "High",
      avg_days_overdue > 30 | credit_utilization > 0.7 ~ "Medium",
      TRUE ~ "Low"
    )
  ) |>
  _______(desc(total_ar)) 

# Create interactive customer table
datatable(
  customer_summary |>
    select(
      Customer = customer_name,
      `Total AR` = total_ar,
      `Credit Limit` = credit_limit,
      `Utilization %` = credit_utilization,
      `Invoices` = invoice_count,
      `Avg Days Overdue` = avg_days_overdue,
      `Risk` = risk_score
    ),
  options = list(
    dom = 'ftip',
    order = list(list(1, 'desc'))
  ),
  rownames = FALSE
) |>
  formatCurrency(c("Total AR", "Credit Limit"), digits = 0) |>
  formatPercentage("Utilization %", digits = 0) |>
  formatStyle(
    "Risk",
    backgroundColor = styleEqual(
      c("Low", "Medium", "High"),
      c("#C8E6C9", "#FFF3E0", "#FFCDD2")
    )
  )
```

## Row

#### Customer Risk Matrix

```{r}
#| label: customer-risk-matrix
# Create risk matrix scatter plot

risk_plot <- customer_summary|>
  ggplot(aes(
    x = avg_days_overdue, 
    y = total_ar,
    color = risk_score,
    size = credit_utilization,
    text = paste(
      "Customer:", customer_name,
      "<br>Total AR:", dollar(total_ar),
      "<br>Avg Days Overdue:", avg_days_overdue,
      "<br>Utilization:", percent(credit_utilization)
    )
  )) +
  geom_point(alpha = 0.7) +
  scale_y_continuous(labels = dollar_format()) +
  scale_color_manual(
    values = c(
      "Low" = ar_colors[["current"]],
      "Medium" = ar_colors[["early"]],
      "High" = ar_colors[["critical"]]
    )
  ) +
  scale_size_continuous(range = c(3, 10)) +
  labs(
    title = "Customer Risk Matrix",
    x = "Average Days Overdue",
    y = "Total AR Outstanding",
    color = "Risk Level",
    size = "Credit\nUtilization"
  )

# Convert to interactive
ggplotly(risk_plot, tooltip = "text") |>
  layout(hovermode = "closest")
```

## Row

#### Credit Utilization Analysis

```{r}
#| label: credit-utilization
# Create credit utilization chart
utilization_data <- customer_summary |>
  mutate(
    utilization_category = case_when(
      credit_utilization >= 1 ~ "Over Limit",
      credit_utilization >= 0.9 ~ "90-100%",
      credit_utilization >= 0.75 ~ "75-90%",
      credit_utilization >= 0.5 ~ "50-75%",
      TRUE ~ "Under 50%"
    ),
    utilization_category = factor(
      utilization_category,
      levels = c("Under 50%", "50-75%", "75-90%", "90-100%", "Over Limit")
    )
  )

utilization_plot <- utilization_data |>
  _______(aes(y = utilization_category)) +
  geom_bar(fill = ar_colors["positive"]) +
  scale_x_continuous(breaks = seq(0, 20, 2)) +
  labs(
    title = "Credit Utilization Distribution",
    y = "Utilization Range",
    x = "Number of Customers"
  ) 

# Convert to interactive
ggplotly(utilization_plot) |>
  layout(hovermode = "x")
```

# Collection Trends

## Row


#### Collection Performance Over Time

```{r}
#| label: collection-trends
# Calculate monthly collection metrics
monthly_collections <- ar_data |>
  filter(!is.na(payment_date)) |>
  mutate(
    payment_month = floor_date(payment_date, "month"),
    days_to_collect = as.numeric(payment_date - invoice_date)
  ) |>
  group_by(payment_month) |>
  summarise(
    collected_amount = sum(amount),
    avg_collection_days = mean(days_to_collect, na.rm = TRUE),
    invoices_collected = n(),
    .groups = "drop"
  ) |>
  filter(payment_month >= today() - months(6))

# Create collection trend plot
collection_plot <- monthly_collections |>
  ggplot(aes(x = payment_month)) +
  geom_col(aes(y = collected_amount), fill = ar_colors["positive"], alpha = 0.7) +
  geom_line(aes(y = avg_collection_days * 5000), color = ar_colors["negative"], linewidth = 2) +
  geom_point(aes(y = avg_collection_days * 5000), color = ar_colors["negative"], size = 3) +
  scale_y_continuous(
    labels = dollar_format(),
    sec.axis = sec_axis(~ . / 5000, name = "Avg Days to Collect")
  ) +
  scale_x_date(date_labels = "%b %Y") +
  _______(
    title = "Monthly Collections and Average Days to Collect",
    x = NULL,
    y = "Amount Collected"
  )

# Convert to interactive
ggplotly(collection_plot) |>
  layout(hovermode = "x unified")
```

## Row
#### Payment Behavior Analysis

```{r}
#| label: payment-behavior
# Analyze payment patterns
payment_patterns <- ar_data |>
  _______(status == "Paid" & !is.na(payment_date)) |>
  mutate(
    days_to_pay = as.numeric(payment_date - invoice_date),
    payment_timing = case_when(
      days_to_pay <= terms_days ~ "On Time",
      days_to_pay <= terms_days + 10 ~ "Slightly Late",
      days_to_pay <= terms_days + 30 ~ "Late",
      TRUE ~ "Very Late"
    ),
    payment_timing = factor(
      payment_timing,
      levels = c("On Time", "Slightly Late", "Late", "Very Late")
    )
  )

# Create payment timing distribution
timing_plot <- payment_patterns |>
  count(payment_timing) |>
  ggplot(_______(x = payment_timing, y = n, fill = payment_timing)) +
  geom_col() +
  scale_fill_manual(
    values = c(
      "On Time" = ar_colors[["current"]],
      "Slightly Late" = ar_colors[["early"]],
      "Late" = ar_colors[["moderate"]],
      "Very Late" = ar_colors[["critical"]]
    )
  ) +
  labs(
    title = "Payment Timing Distribution",
    x = NULL,
    y = "Number of Invoices",
    fill = NULL
  ) 

# Convert to interactive
_______(timing_plot) |>
  layout(hovermode = "x")
```

# Cash Flow Forecast

## Row

### Expected Collections Forecast

```{r}
#| label: collection-forecast
# Create collection forecast based on aging and historical patterns
forecast_data <- ar_aging |>
  _______(
    # Assign collection probability based on aging
    collection_probability = case_when(
      aging_bucket == "Current" ~ 0.95,
      aging_bucket == "1-30 Days" ~ 0.85,
      aging_bucket == "31-60 Days" ~ 0.70,
      aging_bucket == "61-90 Days" ~ 0.50,
      TRUE ~ 0.25
    ),
    # Estimate collection timing
    expected_collection_days = case_when(
      aging_bucket == "Current" ~ 15,
      aging_bucket == "1-30 Days" ~ 20,
      aging_bucket == "31-60 Days" ~ 30,
      aging_bucket == "61-90 Days" ~ 45,
      TRUE ~ 90
    ),
    expected_collection_date = current_date + days(expected_collection_days),
    expected_collection_week = floor_date(expected_collection_date, "week"),
    expected_amount = amount * collection_probability
  )

# Aggregate by week
weekly_forecast <- forecast_data |>
  _______(expected_collection_week <= current_date + weeks(8)) |>
  group_by(expected_collection_week) |>
  summarise(
    expected_collections = sum(expected_amount),
    best_case = sum(amount),
    worst_case = sum(expected_amount * 0.7),
    .groups = "drop"
  )

# Create forecast visualization (static plot - ribbon doesn't convert well to plotly)
weekly_forecast |>
  _______(aes(x = expected_collection_week)) +
  geom_ribbon(
    aes(ymin = worst_case, ymax = best_case),
    fill = ar_colors["positive"],
    alpha = 0.2
  ) +
  geom_line(
    aes(y = expected_collections),
    color = ar_colors["positive"],
    linewidth = 2
  ) +
  geom_point(
    aes(y = expected_collections),
    color = ar_colors["positive"],
    size = 3
  ) +
  scale_y_continuous(labels = dollar_format()) +
  scale_x_date(date_labels = "%b %d") +
  labs(
    title = "8-Week Collection Forecast",
    subtitle = "Expected collections with confidence band",
    x = NULL,
    y = "Expected Collections"
  )
```

## Row

### Collection Probability by Aging

```{r}
#| label: collection-probability-table
# Create collection probability analysis
probability_summary <- ar_aging |>
  _______(aging_bucket) |>
  _______(
    outstanding_amount = sum(amount),
    invoice_count = n(),
    .groups = "drop"
  ) |>
  _______(
    collection_probability = case_when(
      aging_bucket == "Current" ~ 95,
      aging_bucket == "1-30 Days" ~ 85,
      aging_bucket == "31-60 Days" ~ 70,
      aging_bucket == "61-90 Days" ~ 50,
      TRUE ~ 25
    ),
    expected_collection = outstanding_amount * (collection_probability / 100),
    expected_loss = outstanding_amount - expected_collection
  )

# Display probability table
probability_summary |>
  _______() |>
  tab_header(
    title = "Collection Probability Analysis",
    subtitle = "Expected collections based on historical patterns"
  ) |>
  fmt_currency(
    columns = c(outstanding_amount, expected_collection, expected_loss),
    decimals = 0
  ) |>
  fmt_number(
    columns = invoice_count,
    decimals = 0
  ) |>
  fmt_percent(
    columns = collection_probability,
    scale_values = FALSE,
    decimals = 0
  ) |>
  cols_label(
    aging_bucket = "Aging Bucket",
    outstanding_amount = "Outstanding",
    invoice_count = "Invoices",
    collection_probability = "Collection %",
    expected_collection = "Expected",
    expected_loss = "Risk"
  ) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(
      columns = expected_loss,
      rows = expected_loss > 10000
    )
  ) |>
  grand_summary_rows(
    columns = c(outstanding_amount, expected_collection, expected_loss),
    fns = list(Total = ~sum(.)),
    fmt = ~fmt_currency(., decimals = 0)
  )
```
