---
title: "10: Accounts Receivable Dashboard - Solution"
format: dashboard
execute: 
  eval: true
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false

# Load required packages
library(tidyverse)   # for data manipulation
library(scales)      # for formatting numbers
library(DT)          # for interactive tables
library(plotly)      # for interactive plots
library(gt)          # for creating beautiful tables

# Set theme for consistent plot appearance
theme_set(theme_minimal(base_size = 12))

# Define color scheme for AR status
ar_colors <- c(
  current = "#2E7D32",      # Green for current
  early = "#FFC107",        # Yellow for 1-30 days
  moderate = "#FF9800",     # Orange for 31-60 days
  late = "#F44336",         # Red for 61-90 days
  critical = "#B71C1C",     # Dark red for over 90 days
  positive = "#1976D2",     # Blue for positive indicators
  negative = "#C62828"      # Red for negative indicators
)

set.seed(123)

# Create customer data
customers <- tribble(
  ~customer_id, ~customer_name, ~credit_limit, ~industry, ~risk_category,
  "C001", "TechCorp Solutions", 100000, "Technology", "Low",
  "C002", "Global Manufacturing", 150000, "Manufacturing", "Medium",
  "C003", "Retail Express", 75000, "Retail", "High",
  "C004", "Healthcare Plus", 200000, "Healthcare", "Low",
  "C005", "Finance Partners", 125000, "Financial Services", "Medium",
  "C006", "Energy Systems", 175000, "Energy", "Low",
  "C007", "Consumer Goods Co", 80000, "Consumer Goods", "High",
  "C008", "Logistics Network", 90000, "Transportation", "Medium",
  "C009", "Education Services", 60000, "Education", "Low",
  "C010", "Real Estate Group", 110000, "Real Estate", "Medium"
)

# Create invoice data
n_invoices <- 150
invoice_dates <- seq(from = today() - days(180), to = today() - days(1), by = "day")

ar_data <- tibble(
  invoice_id = paste0("INV-", str_pad(1:n_invoices, 5, pad = "0")),
  customer_id = sample(customers$customer_id, n_invoices, replace = TRUE, 
                      prob = c(0.15, 0.12, 0.08, 0.13, 0.10, 0.12, 0.07, 0.08, 0.08, 0.07)),
  invoice_date = sample(invoice_dates, n_invoices, replace = TRUE),
  amount = round(runif(n_invoices, min = 5000, max = 50000), -2),
  terms_days = sample(c(30, 45, 60), n_invoices, replace = TRUE, prob = c(0.7, 0.2, 0.1))
) |>
  mutate(
    due_date = invoice_date + days(terms_days),
    # 70% paid, 30% open
    status = sample(c("Paid", "Open"), n_invoices, replace = TRUE, prob = c(0.7, 0.3)),
    # For paid invoices, create payment dates
    payment_date = if_else(
      status == "Paid",
      invoice_date + days(round(rnorm(n(), mean = terms_days + 5, sd = 15))),
      as.Date(NA)
    ),
    # Ensure payment date is not in the future
    payment_date = if_else(payment_date > today(), as.Date(NA), payment_date),
    status = if_else(is.na(payment_date) & status == "Paid", "Open", status)
  ) |>
  left_join(customers, by = "customer_id")

# Save for student use
# write_rds(ar_data, "data/10-assignment-ar_data.rds")
```

# Executive Summary {.tabset}

As of January 31, 2026, the company faces significant collection risk with $1.13M (59%) of the $1.9M total accounts receivable past due, including $485,700 severely delinquent beyond 90 days concentrated primarily in the Technology and Transportation sectors. Nine of ten major customers are classified as high-risk, with credit utilization rates ranging from 134% to 279% above their limits—most notably TechCorp Solutions and Retail Express—signaling an urgent need to reassess credit policies and enforce collection protocols on these accounts. 

## Key Metrics

### Row {height="25%"}

```{r}
#| label: load-data
# Load the AR data (using created sample data)
# ar_data <- read_rds("data/10-assignment-ar_data.rds")

# Set current date for aging calculations
current_date <- ymd("2026-01-31")

```

```{r}
#| label: calculate-kpis
# Calculate key performance indicators

# Total AR balance
total_ar <- ar_data |>
  filter(status == "Open") |>
  summarise(total = sum(amount)) |>
  pull(total)

# Calculate Days Sales Outstanding (DSO)
# Calculate average daily sales from last 90 days of paid invoices
avg_daily_sales <- ar_data |>
  filter(status == "Paid" & payment_date >= today() - days(90)) |>
  summarise(daily_sales = sum(amount) / 90) |>
  pull(daily_sales)

current_dso <- round(total_ar / avg_daily_sales, 1)

# Calculate past due percentage
past_due_amount <- ar_data |>
  filter(status == "Open" & due_date < current_date) |>
  summarise(total = sum(amount)) |>
  pull(total)

past_due_percentage <- past_due_amount / total_ar

# Count of overdue invoices
overdue_count <- ar_data |>
  filter(status == "Open" & due_date < current_date) |>
  nrow()

# Average days to pay (for paid invoices)
avg_days_to_pay <- ar_data |>
  filter(status == "Paid" & !is.na(payment_date)) |>
  mutate(days_to_pay = as.numeric(payment_date - invoice_date)) |>
  summarise(avg_days = mean(days_to_pay, na.rm = TRUE)) |>
  pull(avg_days)
```

### Row {height="75%"}

## Summary Statistics

```{r}
#| label: summary-table
# Create comprehensive AR summary statistics
summary_stats <- ar_data |>
  filter(status == "Open") |>
  summarise(
    `Total Outstanding AR` = sum(amount),
    `Number of Open Invoices` = n(),
    `Average Invoice Amount` = mean(amount),
    `Median Invoice Amount` = median(amount),
    `Largest Invoice` = max(amount),
    `Smallest Invoice` = min(amount),
    `Total Past Due` = sum(amount[due_date < current_date]),
    `Current (Not Due)` = sum(amount[due_date >= current_date])
  )

# Create summary table using gt
summary_stats |>
  pivot_longer(everything(), names_to = "Metric", values_to = "Value") |>
  gt() |>
  tab_header(
    title = "Accounts Receivable Summary",
    subtitle = paste("As of", format(current_date, "%B %d, %Y"))
  ) |>
  fmt_currency(
    columns = Value,
    rows = c(1, 3:8),
    decimals = 0
  ) |>
  fmt_number(
    columns = Value,
    rows = 2,
    decimals = 0
  ) |>
  cols_label(
    Metric = "Metric",
    Value = "Value"
  ) |>
  tab_style(
    style = cell_fill(color = "#E3F2FD"),
    locations = cells_body(rows = c(1, 7))
  )
```

# Aging Analysis

## Row {height="30%"}

### Aging Summary

```{r}
#| label: create-aging-buckets
# Create aging buckets for AR analysis
ar_aging <- ar_data |>
  filter(status == "Open") |>
  mutate(
    # Calculate days outstanding
    days_outstanding = as.numeric(current_date - invoice_date),
    # Calculate days overdue
    days_overdue = as.numeric(current_date - due_date),
    days_overdue = if_else(days_overdue < 0, 0, days_overdue),
    # Create aging buckets
    aging_bucket = case_when(
      days_overdue == 0 ~ "Current",
      days_overdue <= 30 ~ "1-30 Days",
      days_overdue <= 60 ~ "31-60 Days",
      days_overdue <= 90 ~ "61-90 Days",
      TRUE ~ "Over 90 Days"
    ),
    # Set factor levels for proper ordering
    aging_bucket = factor(aging_bucket, 
                         levels = c("Current", "1-30 Days", "31-60 Days", 
                                   "61-90 Days", "Over 90 Days"))
  )

# Create aging summary table
aging_summary <- ar_aging |>
  group_by(aging_bucket) |>
  summarise(
    `Invoice Count` = n(),
    `Total Amount` = sum(amount),
    `Percentage` = sum(amount) / sum(ar_aging$amount),
    .groups = "drop"
  )

# Display aging summary with gt
aging_summary |>
  gt() |>
  tab_header(
    title = "AR Aging Summary",
    subtitle = "Distribution by aging bucket"
  ) |>
  fmt_number(
    columns = `Invoice Count`,
    decimals = 0
  ) |>
  fmt_currency(
    columns = `Total Amount`,
    decimals = 0
  ) |>
  fmt_percent(
    columns = Percentage,
    decimals = 0
  ) |>
  cols_label(
    aging_bucket = "Aging Bucket"
  ) |>
  # Apply conditional formatting based on aging
  tab_style(
    style = cell_fill(color = "#C8E6C9"),
    locations = cells_body(
      columns = everything(),
      rows = aging_bucket == "Current"
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#FFCDD2"),
    locations = cells_body(
      columns = everything(),
      rows = aging_bucket == "Over 90 Days"
    )
  )
```

## Row {height="35%"}


#### Aging Distribution 

```{r}
#| label: aging-distribution-plot
# Create aging distribution bar chart
aging_plot <- ar_aging |>
  group_by(aging_bucket) |>
  summarise(total = sum(amount), .groups = "drop") |>
  ggplot(aes(x = aging_bucket, y = total, fill = aging_bucket)) +
  geom_col() +
  scale_y_continuous(labels = dollar_format()) +
  scale_fill_manual(
    values = c(
      "Current" = ar_colors[["current"]],
      "1-30 Days" = ar_colors[["early"]],
      "31-60 Days" = ar_colors[["moderate"]],
      "61-90 Days" = ar_colors[["late"]],
      "Over 90 Days" = ar_colors[["critical"]]
    )
  ) +
  labs(
    title = "AR Aging Distribution",
    x = NULL,
    y = "Amount Outstanding",
    fill = NULL
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# Convert to interactive plot
ggplotly(aging_plot, tooltip = c("x", "y")) |>
  layout(hovermode = "x unified")
```

## Row {height="35%"}

#### Aging by Customer Type

```{r}
#| label: aging-by-customer-type
# Create aging analysis by customer industry/type
industry_aging <- ar_aging |>
  group_by(industry, aging_bucket) |>
  summarise(total = sum(amount), .groups = "drop")

# Create stacked bar chart - HORIZONTAL VERSION
industry_plot <- industry_aging |>
  ggplot(aes(y = reorder(industry, total, sum), x = total, fill = aging_bucket)) +  # Swapped x and y, added sum function
  geom_col(position = "stack") +
  scale_x_continuous(labels = dollar_format()) +  # Changed to scale_x
  scale_fill_manual(
    values = c(
      "Current" = ar_colors[["current"]],
      "1-30 Days" = ar_colors[["early"]],
      "31-60 Days" = ar_colors[["moderate"]],
      "61-90 Days" = ar_colors[["late"]],
      "Over 90 Days" = ar_colors[["critical"]]
    )
  ) +
  labs(
    title = "Aging by Industry",
    y = NULL,  # Swapped labels
    x = "Amount Outstanding",
    fill = NULL
  )

# Convert to interactive
ggplotly(industry_plot) |>
  layout(hovermode = "y unified")  # Changed to y unified
```

# Customer Analysis

## Row

### Top Customers by Exposure

```{r}
#| label: top-customers-table
# Calculate customer exposure metrics
customer_summary <- ar_aging |>
  group_by(customer_id, customer_name, credit_limit) |>
  summarise(
    total_ar = sum(amount),
    invoice_count = n(),
    oldest_invoice = min(invoice_date),
    days_oldest = as.numeric(current_date - oldest_invoice),
    avg_days_overdue = round(mean(days_overdue, na.rm = TRUE), 0),
    .groups = "drop"
  ) |>
  mutate(
    credit_utilization = total_ar / credit_limit,
    risk_score = case_when(
      avg_days_overdue > 60 | credit_utilization > 0.9 ~ "High",
      avg_days_overdue > 30 | credit_utilization > 0.7 ~ "Medium",
      TRUE ~ "Low"
    )
  ) |>
  arrange(desc(total_ar)) 

# Create interactive customer table
datatable(
  customer_summary |>
    select(
      Customer = customer_name,
      `Total AR` = total_ar,
      `Credit Limit` = credit_limit,
      `Utilization %` = credit_utilization,
      `Invoices` = invoice_count,
      `Avg Days Overdue` = avg_days_overdue,
      `Risk` = risk_score
    ),
  options = list(
    pageLength = 10,
    dom = 'ftip',
    order = list(list(1, 'desc'))
  ),
  rownames = FALSE
) |>
  formatCurrency(c("Total AR", "Credit Limit"), digits = 0) |>
  formatPercentage("Utilization %", digits = 0) |>
  formatStyle(
    "Risk",
    backgroundColor = styleEqual(
      c("Low", "Medium", "High"),
      c("#C8E6C9", "#FFF3E0", "#FFCDD2")
    )
  )
```

## Row

### Column {width="50%"}

#### Customer Risk Matrix

```{r}
#| label: customer-risk-matrix
# Create risk matrix scatter plot
risk_data <- customer_summary |>
  slice_head(n = 20)

risk_plot <- risk_data |>
  ggplot(aes(
    x = avg_days_overdue, 
    y = total_ar,
    color = risk_score,
    size = credit_utilization,
    text = paste(
      "Customer:", customer_name,
      "<br>Total AR:", dollar(total_ar),
      "<br>Avg Days Overdue:", avg_days_overdue,
      "<br>Utilization:", percent(credit_utilization)
    )
  )) +
  geom_point(alpha = 0.7) +
  scale_y_continuous(labels = dollar_format()) +
  scale_color_manual(
    values = c(
      "Low" = ar_colors[["current"]],
      "Medium" = ar_colors[["early"]],
      "High" = ar_colors[["critical"]]
    )
  ) +
  scale_size_continuous(range = c(3, 10)) +
  labs(
    title = "Customer Risk Matrix",
    x = "Average Days Overdue",
    y = "Total AR Outstanding",
    color = "Risk Level",
    size = "Credit\nUtilization"
  )

# Convert to interactive
ggplotly(risk_plot, tooltip = "text") |>
  layout(hovermode = "closest")
```

### Column {width="50%"}

#### Credit Utilization Analysis

```{r}
#| label: credit-utilization
# Create credit utilization chart
utilization_data <- customer_summary |>
  mutate(
    utilization_category = case_when(
      credit_utilization >= 1 ~ "Over Limit",
      credit_utilization >= 0.9 ~ "90-100%",
      credit_utilization >= 0.75 ~ "75-90%",
      credit_utilization >= 0.5 ~ "50-75%",
      TRUE ~ "Under 50%"
    ),
    utilization_category = factor(
      utilization_category,
      levels = c("Under 50%", "50-75%", "75-90%", "90-100%", "Over Limit")
    )
  )

utilization_plot <- utilization_data |>
  ggplot(aes(y = utilization_category)) +
  geom_bar(fill = ar_colors["positive"]) +
  scale_x_continuous(breaks = seq(0, 20, 2)) +
  labs(
    title = "Credit Utilization Distribution",
    y = "Utilization Range",
    x = "Number of Customers"
  ) 

# Convert to interactive
ggplotly(utilization_plot) |>
  layout(hovermode = "x")
```

# Collection Trends

## Row


#### Collection Performance Over Time

```{r}
#| label: collection-trends
# Calculate monthly collection metrics
monthly_collections <- ar_data |>
  filter(!is.na(payment_date)) |>
  mutate(
    payment_month = floor_date(payment_date, "month"),
    days_to_collect = as.numeric(payment_date - invoice_date)
  ) |>
  group_by(payment_month) |>
  summarise(
    collected_amount = sum(amount),
    avg_collection_days = mean(days_to_collect, na.rm = TRUE),
    invoices_collected = n(),
    .groups = "drop"
  ) |>
  filter(payment_month >= today() - months(6))

# Create collection trend plot
collection_plot <- monthly_collections |>
  ggplot(aes(x = payment_month)) +
  geom_col(aes(y = collected_amount), fill = ar_colors["positive"], alpha = 0.7) +
  geom_line(aes(y = avg_collection_days * 5000), color = ar_colors["negative"], linewidth = 2) +
  geom_point(aes(y = avg_collection_days * 5000), color = ar_colors["negative"], size = 3) +
  scale_y_continuous(
    labels = dollar_format(),
    sec.axis = sec_axis(~ . / 5000, name = "Avg Days to Collect")
  ) +
  scale_x_date(date_labels = "%b %Y") +
  labs(
    title = "Monthly Collections and Average Days to Collect",
    x = NULL,
    y = "Amount Collected"
  )

# Convert to interactive
ggplotly(collection_plot) |>
  layout(hovermode = "x unified")
```

## Row
#### Payment Behavior Analysis

```{r}
#| label: payment-behavior
# Analyze payment patterns
payment_patterns <- ar_data |>
  filter(status == "Paid" & !is.na(payment_date)) |>
  mutate(
    days_to_pay = as.numeric(payment_date - invoice_date),
    payment_timing = case_when(
      days_to_pay <= terms_days ~ "On Time",
      days_to_pay <= terms_days + 10 ~ "Slightly Late",
      days_to_pay <= terms_days + 30 ~ "Late",
      TRUE ~ "Very Late"
    ),
    payment_timing = factor(
      payment_timing,
      levels = c("On Time", "Slightly Late", "Late", "Very Late")
    )
  )

# Create payment timing distribution
timing_plot <- payment_patterns |>
  count(payment_timing) |>
  ggplot(aes(x = payment_timing, y = n, fill = payment_timing)) +
  geom_col() +
  scale_fill_manual(
    values = c(
      "On Time" = ar_colors[["current"]],
      "Slightly Late" = ar_colors[["early"]],
      "Late" = ar_colors[["moderate"]],
      "Very Late" = ar_colors[["critical"]]
    )
  ) +
  labs(
    title = "Payment Timing Distribution",
    x = NULL,
    y = "Number of Invoices",
    fill = NULL
  ) 

# Convert to interactive
ggplotly(timing_plot) |>
  layout(hovermode = "x")
```

# Cash Flow Forecast

## Row

### Expected Collections Forecast

```{r}
#| label: collection-forecast
# Create collection forecast based on aging and historical patterns
forecast_data <- ar_aging |>
  mutate(
    # Assign collection probability based on aging
    collection_probability = case_when(
      aging_bucket == "Current" ~ 0.95,
      aging_bucket == "1-30 Days" ~ 0.85,
      aging_bucket == "31-60 Days" ~ 0.70,
      aging_bucket == "61-90 Days" ~ 0.50,
      TRUE ~ 0.25
    ),
    # Estimate collection timing
    expected_collection_days = case_when(
      aging_bucket == "Current" ~ 15,
      aging_bucket == "1-30 Days" ~ 20,
      aging_bucket == "31-60 Days" ~ 30,
      aging_bucket == "61-90 Days" ~ 45,
      TRUE ~ 90
    ),
    expected_collection_date = current_date + days(expected_collection_days),
    expected_collection_week = floor_date(expected_collection_date, "week"),
    expected_amount = amount * collection_probability
  )

# Aggregate by week
weekly_forecast <- forecast_data |>
  filter(expected_collection_week <= current_date + weeks(8)) |>
  group_by(expected_collection_week) |>
  summarise(
    expected_collections = sum(expected_amount),
    best_case = sum(amount),
    worst_case = sum(expected_amount * 0.7),
    .groups = "drop"
  )

# Create forecast visualization (static plot - ribbon doesn't convert well to plotly)
weekly_forecast |>
  ggplot(aes(x = expected_collection_week)) +
  geom_ribbon(
    aes(ymin = worst_case, ymax = best_case),
    fill = ar_colors["positive"],
    alpha = 0.2
  ) +
  geom_line(
    aes(y = expected_collections),
    color = ar_colors["positive"],
    linewidth = 2
  ) +
  geom_point(
    aes(y = expected_collections),
    color = ar_colors["positive"],
    size = 3
  ) +
  scale_y_continuous(labels = dollar_format()) +
  scale_x_date(date_labels = "%b %d") +
  labs(
    title = "8-Week Collection Forecast",
    subtitle = "Expected collections with confidence band",
    x = NULL,
    y = "Expected Collections"
  )
```

## Row

### Collection Probability by Aging

```{r}
#| label: collection-probability-table
# Create collection probability analysis
probability_summary <- ar_aging |>
  group_by(aging_bucket) |>
  summarise(
    outstanding_amount = sum(amount),
    invoice_count = n(),
    .groups = "drop"
  ) |>
  mutate(
    collection_probability = case_when(
      aging_bucket == "Current" ~ 95,
      aging_bucket == "1-30 Days" ~ 85,
      aging_bucket == "31-60 Days" ~ 70,
      aging_bucket == "61-90 Days" ~ 50,
      TRUE ~ 25
    ),
    expected_collection = outstanding_amount * (collection_probability / 100),
    expected_loss = outstanding_amount - expected_collection
  )

# Display probability table
probability_summary |>
  gt() |>
  tab_header(
    title = "Collection Probability Analysis",
    subtitle = "Expected collections based on historical patterns"
  ) |>
  fmt_currency(
    columns = c(outstanding_amount, expected_collection, expected_loss),
    decimals = 0
  ) |>
  fmt_number(
    columns = invoice_count,
    decimals = 0
  ) |>
  fmt_percent(
    columns = collection_probability,
    scale_values = FALSE,
    decimals = 0
  ) |>
  cols_label(
    aging_bucket = "Aging Bucket",
    outstanding_amount = "Outstanding",
    invoice_count = "Invoices",
    collection_probability = "Collection %",
    expected_collection = "Expected",
    expected_loss = "Risk"
  ) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(
      columns = expected_loss,
      rows = expected_loss > 10000
    )
  ) |>
  grand_summary_rows(
    columns = c(outstanding_amount, expected_collection, expected_loss),
    fns = list(Total = ~sum(.)),
    fmt = ~fmt_currency(., decimals = 0)
  )
```
